---
title: "Tidy Data"
subtitle: "5 Most Common Problems With Messy Datasets"
author: "Pengfei Qin"
date: "12 April 2021"
output:
  xaringan::moon_reader:
    css: ['default', 'metropolis', 'metropolis-fonts', 'my-css.css']
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      
---

<style>
.remark-slide-content.hljs-default {
  border-top: 40px solid #23373B;
}

.remark-slide-content > h1 {
  font-size: 26px;
  margin-top: -80px;
}

</style>



```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(eval = TRUE) ## Show all R output
library(pacman)
p_load(tidyverse, dplyr, tidyr, modeldata, skimr, tidymodels, data.table, knitr, stargazer)
```


# Why we need data tidying?


--


### Problems



- The structure of the data set is not uniform (various arrangements in variables and observations).

--


### Solution

- Data Tidying: Structuring datasets to facilitate analysis.

- Tidy data provides us a standard method to organize the values in a dataset, and it can link the structure of a dataset with its semantics.

--

### Purpose

- Tidying the initial data can make the followinghu data analysis easier.

--

- Consistent structured datasets will also make our data analysis easier.



---
# The principle of tidy data


### Tidy data

1. Every column is a variable.

2. Every row is an observation.

3. Each type of observational unit forms a table.

--

```{r}
dataset4 = data.frame(Song = rep(c("Song_A", "Song_B", "Song_C"), each =3),
  'Year' = rep(c("2019", "2020", "2021"), each = 3),
  'Time' = rep(c("241", "175", "239"), each = 3),
  'Date'= c("2021-04-01", "2021-04-02", "2021-04-03", "2021-04-01", "2021-04-02", "2021-04-03", "2021-04-01", "2021-04-02", "2021-04-03"),
  'Rank' = c(1,3,2,2,1,3,3,2,1))
dataset4
```

---
class: inverse, center, middle
name: common problems

# "5 Most Common Problems With Messy Datasets"

---
# 1. Column headers are values, not variable names.

$~$

```{r}
dataset1 = data.frame(name = c("Company_A", "Company_B", "Company_C", "Company_D"),
  'bachelor'= sample(1:10, 4),
  'master'= sample(1:10, 4),
  'PhD'= sample(1:10, 4))
dataset1
```


---
# 1. Column headers are values, not variable names.

$~$

```{r}
dataset1 <- dataset1 %>% 
  pivot_longer(-name, names_to = "degree", values_to = "frequency")
dataset1
```



---
# 2. Multiple variables stored in one column

$~$

```{r}
dataset2 = data.frame(name = c("Company_A", "Company_B", "Company_C", "Company_D"),
  'm2035'= sample(1:10, 4),
  'm3550'= sample(1:10, 4),
  'm5065'= sample(1:10, 4),
  'f2035'= sample(1:10, 4),
  'f3550'= sample(1:10, 4),
  'f5060'= sample(1:10, 4))
dataset2
```

---
# 2. Multiple variables stored in one column

$~$

```{r}
dataset2 <- dataset2 %>% 
  pivot_longer(-name, names_to = "combination", values_to = "frequency")
dataset2
```

---
# 2. Multiple variables stored in one column

$~$
```{r}
dataset2 <- dataset2 %>% 
  separate(combination, c("sex", "age"),sep = "(?<=[a-z])(?=[0-9])")
```


```{r, echo = FALSE}
dataset2
```

---


# 3. Variables are stored in both columns and rows.

</br>

```{r}
dataset3 = data.frame(city = rep(c("Beijing", "Hong Kong", "Los Angeles", "New York"), each =2),
  'month'= c("January"),
  'element'= c("avg_environmental_quality", "avg_air_quality"),
  'y2019'= sample(c("high", "median", "low"),8, replace = TRUE),
  'y2020'= sample(c("high", "median", "low"),8, replace = TRUE),
  'y2021'= sample(c("high", "median", "low"),8, replace = TRUE))
dataset3
```

---

# 3. Variables are stored in both columns and rows.

</br>

```{r}
dataset3 <- dataset3 %>% 
  pivot_longer(y2019:y2021, names_to = "year", values_to = "value")

dataset3 <- dataset3 %>% 
  mutate(year = as.integer(gsub("y", "", year))) %>%
  select(city, month, element, year, value)

dataset3
```

---

# 3. Variables are stored in both columns and rows.

</br>

```{r}
dataset3 <- dataset3 %>% 
  pivot_wider(names_from = element, values_from = value)
dataset3
```


---

# 4. Multiple types of observational units are stored in the same table.

</br>

```{r, echo=FALSE}
dataset4
```

---

# 4. Multiple types of observational units are stored in the same table.

```{r}
song <- dataset4 %>% 
  distinct(Song, Year, Time)
song
```


```{r}
rank <- dataset4 %>%
  distinct(Song, Date, Rank)
rank
```


---
# 5. A single observational unit is stored in multiple tables.

--



```{r}
GDP_and_Tax = data.frame(City = c("Beijing", "Hong Kong", "New York", "Portland", "Eugene"),
  'GDP'= runif(5, 100, 200),
  'Tax_Revenue' = runif(5, 15, 25))
GDP_and_Tax
```

```{r}
Energy_and_Industry = data.frame(City = c("Beijing", "Hong Kong", "New York", "Portland", "Eugene"),
  'Energy_Consumption'= runif(5, 1900, 2000),
  'Industrial_Output' = runif(5, 250, 300))
Energy_and_Industry
```

---

# 5. A single observational unit is stored in multiple tables.

</br>

```{r}
dataset5 <- inner_join(GDP_and_Tax, Energy_and_Industry)
dataset5
```


